---
title: "Identifying partners at sea from joint movement metrics of pelagic pair trawlers"
author: 
  - Rocío Joo:
      institute: [IFREMER,UF,GFW]
  - Nicolas Bez:
      institute: MARBEC
  - Marie-Pierre Etienne:
      institute: Rennes
  - Pablo Marin:
      institute: IMARPE
  - Nicolas Goascoz:
      institute: IFREMER2
  - Jérôme Roux:
      institute: IFREMER
  - Stéphanie Mahévas:
      institute: IFREMER
      
institute:
  - IFREMER: IFREMER, Ecologie et Modèles pour l'Halieutique, BP 21105, 44311 Nantes Cedex 03, France
  - UF: Department of Wildlife Ecology and Conservation, Fort Lauderdale Research and Education Center, University of Florida, Fort Lauderdale, FL, USA
  - GFW: Global Fishing Watch, Washington, DC 20036, USA
  - MARBEC: MARBEC, Univ Montpellier, IRD, Ifremer, CNRS, Sète, France
  - Rennes: Univ Rennes, Agrocampus Ouest, CNRS, IRMAR - UMR 6625, F-35000 Rennes, France
  - IMARPE: Instituto del Mar del Perú (IMARPE), Chucuito, Callao, Peru
  - IFREMER2: IFREMER, Laboratoire Ressources Halieutiques de Port en Bessin, BP 32, 14520 Port en Bessin, France
  
      
date: ''
output:
  word_document:
    df_print: paged
    fig_caption: yes
    reference_docx: doc-style.docx
    pandoc_args:
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '2'
  tufte::tufte_html:
    toc: yes
    toc_depth: 2
bibliography: references_total.bib
csl: ices-journal-of-marine-science.csl
---

## Abstract

Here we present an approach to identify partners at sea based on fishing track
analysis, and describe this behaviour in six different fleets: 1) pelagic pair
trawlers, 2) large bottom otter trawlers, 3) small bottom otter trawlers, 4)
mid-water otter trawlers, all operating in the North-East Atlantic Ocean, 5)
anchovy purse-seiners in the South-East Pacific Ocean, and 6) tuna purse-seiners
in the Western Indian Ocean. This type of behaviour is known to exist within
pelagic pair trawlers, since these vessels need to be in pairs for their fishing
operations. However, in practice, some of them decide to move together
throughout their whole fishing trips, and others for only a segment of their
trips. To identify partners at sea, we used a heuristic approach based on
joint-movement metrics computed from vessel monitoring system data and Gaussian
mixture models. The models were fitted to joint-movement metrics of the pelagic
pair trawlers, and subsequently used to identify partners at sea in other
fleets. We found partners at sea in all of the fleets except for  the tuna
purse-seiners. We then analysed the connections at the scale of vessels and
identified exclusive partners. Exclusiveness was more common in pelagic pair
trawlers and small bottom otter trawlers, with $82\%$ and $73\%$ of the vessels
involved in partnerships having exclusive partners. This work shows that there
are collective tactics at least at a pairwise level in diverse fisheries in the
world.

## Keywords

collective behaviour; dyadic joint movement metrics; Gaussian mixture model;
vessel monitoring system; fishing tactics.



# Introduction

Understanding fisher spatial behaviour contributes to the development of
effective spatial management tools. The increasing availability of georeferenced
data from sources like Automatic Identification System (AIS; @Robards2016) and
Vessel Monitoring System (VMS; @Hinz2013) has enabled a proliferation of studies
that characterise fisher spatial dynamics (e.g. @Bertrand2005; @Joo2014),
propose movement models (e.g. @Vermard2010; @Walker2010; @Joo2013;
@Gloaguen2015a), account for it in stock assessment models for fisheries
management (e.g. @Vigier2018) and discuss management measures based on it (e.g.
@Gerritsen2012; @Holmes2011). While individual movement of fishers has been
extensively studied by means of trajectory data, the collective behaviour of
fishermen has been rather neglected. Fishers are social individuals that may
develop collaboration or competing strategies (e.g. @Horta2012; @Hancock1995).
The characterisation of their collective behaviour could provide valuable inputs
that would increase the realism of movement models and make management measures
more effective [@Salas2004; @Rijnsdorp2011; @Gezelius2007].

Collective behaviour can emerge at large or small group scales, and may be
reflected in a variety of movement patterns. Here, we focused on a particular
collective behaviour, which is dyadic or pairwise joint movement behaviour, and
more specifically, aimed at identifying partners at sea, defined as two fishing
vessels that move together at sea. An extensive review and comparison of metrics
for assessing dyadic joint movement [@Joo2018] showed that the metrics varied in
their sensitivity to three aspects of joint movement: proximity, coordination in
direction and coordination in speed. Here, we defined partners at sea as showing
coordinated and proximal joint movement. To account for all of these aspects, we
chose one metric for each of the three dimensions of joint movement, from the
ones recommended in @Joo2018, to characterise the dyadic movement of fishing
vessels.

Strong partnership at sea was expected to be found in pelagic pair trawlers:
since they need to be in pairs at least during each fishing operation, they are
likely to be paring throughout their entire fishing trips. For that reason, in
this study, we aimed at defining the model parameters that would allow us to
identify strong partnership at sea in pelagic pair trawlers in the North-East
Atlantic Ocean through the analysis of their VMS data. After that, the goal was
two-fold: assessing whether if the same patterns of partnership were present in
other fleets; and, if present, assessing the level of exclusiveness in the
partnership within each fleet.

Dyads, or potential candidates for partners at sea, were defined as pairs of
segments of VMS tracks at sea at the same time. For each dyad, three joint
movement metrics were calculated. Then, we fitted a Gaussian mixture model (GMM)
to distinguish three groups of dyads sharing the same types of behaviour. One of
these components was expected to correspond to partners at sea patterns. After
characterising at-sea partnership in this fleet, we used the fitted model to
identify partners at sea in several other fisheries: bottom and mid-water otter
trawlers in the North-East Atlantic Ocean, anchovy purse-seiners in the
South-East Pacific Ocean, and tuna purse-seiners in the Western Indian Ocean. We
showed that this type of behaviour is not exclusive to pelagic pair trawlers,
and discuss possible implications of this behaviour in terms of fishing
strategies. Perspectives opened by this work for further research in collective
spatial behaviour are also discussed.


# Materials and Methods

## Fishing vessels trajectory data

In this section, the VMS data and fishing trip characteristics of the analysed
fleets are briefly described. These are: $1$) French pelagic pair trawlers, $2$)
French large bottom otter trawlers, $3$) French small bottom otter trawlers,
$4$) French mid-water otter trawlers, all operating in the North-East Atlantic
Ocean,  $5$) French tuna purse-seiners in the Western Indian Ocean, and $6$)
Peruvian anchovy purse-seiners in the South-East Pacific Ocean.

For the French fleets, the use of VMS started to be legislated and mandatory in
the European Union since 2000. In practice, records are transmitted at $\sim$ 1
h intervals. In the North-East Atlantic Ocean, we analysed VMS data from fishing
trips performed between 2012 and 2013 within the English Channel and the Celtic
Sea, while in the Indian Ocean, we analysed fishing trips from 2011 to 2013. In
Peru, industrial purse-seiners are also legally obliged to use VMS tracking
devices since 2000, transmitting their positions at $\sim$ 1 h intervals, but
since 2015, VMS positions are recorded each $10$ minutes. We focus on Peruvian
fishing trips during a specific fishing season in 2016.

### French pelagic pair trawlers

A pelagic pair trawl is a gear defined by one trawl towed in midwater by two
vessels to target pelagic fish. Thus, vessels of the pelagic pair trawler fleet
remain close performing almost synchronous movements while operating the trawl.
The distance between vessels during this operation varies between 50 m and 250
m, depending on the warp length (which in turn depends on several factors such
as the fishing depth and technique) [@Prado1988]. The vessels do not need to
move together throughout their whole fishing trips, especially when steaming,
using single trawls or exploring the sea individually looking for shoals
[@Sainsbury1996]. These vessels can spend part of their fishing trips on
individual activities, even targetting other fish that do not require pair
trawling. Most of the pair-trawler fishing trips in the dataset were performed
by relatively large vessels (18-24 m; $\sim 80\%$), and they last $\sim 99h$ on
average, according to fisher logbooks.

### French large and small bottom otter trawlers

The bottom otter trawl gear is a trawl towed by a single vessel; these vessels
target bottom and demersal species. Vessels performing bottom otter trawl
fishing trips had a large variability in their sizes: from 10 to 40 m. The
duration of the trips were proportionally related to the size of the vessels:
larger vessels performed longer trips and generally offshore. Since, for this
type of gear, the spatial behaviour from smaller vessels differs from that of
larger vessels (e.g. the trips are not only shorter but also closer to the
coast), we separated bottom otter trawlers into two groups: one with vessels
smaller than 12 m or performing trips of less than 20 h (we assume that in very
short trips even large vessels act like the small ones), and another one with
vessels larger than 12 m or performing trips of larger duration; vessels with
these characteristics are considered as composing the small otter trawl and
large otter trawl fishing fleets, respectively. The average duration of fishing
trips for both fleets were $\sim$ 16 and $\sim$ 105 hours, respectively,
according to fisher logbooks.

### French mid-water otter trawlers

A mid-water otter trawl gear is also operated by an individual vessel. As the
vessels in the pair trawler fleet, mid-water otter trawlers target pelagic fish
mostly. As with bottom trawlers, vessels performing mid-water trawling trips had
sizes ranging from 10 to 40 m; larger vessels exist (e.g. 90 m long targeting
blue whiting) but were not found in this dataset. However, the spatial behaviour
of these vessels was not conditioned by their size, so they were not separated
by size. The average duration of a fishing trip was $\sim$ 31 hours (fisher
logbooks). Since fishing with mid-water or bottom otter trawls does not require
pair-work, if it exists, it would reflect a strategic/tactical choice.

### French tuna purse-seiners

The fleet is composed of ten to twenty vessels operating in the Indian Ocean and
the size of the purse seiners is typically of sixty meters. Tuna purse-seiners'
fishing trips usually last several tens of days. The time windows targeted in
the present study (2011-2013) followed a harsh period of strong security issues
induced by piracy attacks in the Indian Ocean. During the second half of 2009,
it became mandatory for fishing vessels operating in the Indian Ocean to fish in
pairs before some military protection were enforced. However, some vessels could
have decided to continue moving more or less in pairs as a precautionary
approach. Since tuna purse-seiners perform long fishing trips, we did not expect
vessels to move together throughout their whole fishing trips, but rather over
some shorter opportunist periods of time, eventually changing partners.

### Peruvian anchovy purse-seiners

The ten-minutes frequency of data recording is particularly suiting for
monitoring the anchovy (*Engraulis ringens*) industrial fishery, where
fishing trips usually last less than 24 hours (a median of $17$ hours for the
analysed data), since fish tends to distribute close to the coast in dense
patches [@Bertrand2008; @Joo2014]. In this fishery, vessel size is measured in
terms of its hold capacity, which varies from $32.5$ MT to $900$ MT, with a
median at $\sim 100$ MT. We used data from the first fishing season of 2016 (39
days between June and July). Though the race for fish stopped in 2009 (the total
allowable catch was replaced by an individual vessel quota system; @Aranda2009),
the high abundance of anchovy, the eagerness to save fuel oil and the habit of
performing very short fishing trips, make it common for vessels to go to the
same fishing zones or to follow each other as a fishing tactic. Thus here as
well, we expected to find some patterns of joint movement, although not
perfectly synchronous or remaining close to each other all the time.

	
## Methods		
		
Identifying partners at sea basically consists of 1) data pre-processing and
dyad constitution (i.e. the VMS data was first cleaned and interpolated, and
then dyadic segments of trajectories were identified); 2) joint-movement metrics
derivation for each dyad; 3) identification of clusters of dyadic joint movement
--and particularly partners at sea-- via GMMs; and 4) characterisation of
partnership at vessel and fleet scales. All the analyses were performed in R
[@R2015].

### Data Pre-processing

From the trawler VMS data, fishing trips where at least one pair of consecutive
records were lagged by more than three hours were removed ($\approx 9\%$ of the
total number of fishing trips). For tuna purse-seiners, we used a one-hour
threshold. If there were consecutive records separated for more than one hour,
those differences had to represent less than $10\%$ of the trip duration to keep
the trip in the dataset ($\approx 7\%$ of the total number of fishing trips were
removed). Then, since location records had irregular time steps, we linearly
interpolated tracks to obtain regular 1-hour time steps and simultaneous VMS
positions (i.e. fixes) from vessels at sea. The anchovy purse-seine data was
processed using the SISESATools R package
[https://github.com/PabloMBooster/SISESATools](https://github.com/PabloMBooster/SISESATools)
prior to this study. The SISESATools algorithms apply a two-hour threshold for
consecutive records and use a linear interpolation at 10-minute time steps. From
the (interpolated) fixes, we derived motion variables such as displacement
(distance between consecutive fixes) and absolute angle (between the direction
of the x-axis and the locations at consecutive fixes). The adehabitatLT package
in R [@adehabitatLT] was used to compute those metrics.

We then formed the dyads that would be candidates for partners at sea. Dyads
were defined as the concomitant parts of two vessel tracks crossing each other
at least once during their fishing trips. We considered that, to 'cross each
other', vessels had to be at a proximity of <5 km at least once for all
fleets, except tuna purse-seiners. The latter have a greater range of motion and
do not get so close; for them, the distance threshold was set to 60 km. If both
vessels departed from port and then arrived to port at the same time, the dyad
was to be composed of the two tracks of their whole fishing trips; if not, the
dyad would have been composed by track segments of their fishing trips
corresponding to moments when both vessels were at sea. To keep only dyads with
segments that were long enough for the analysis, an arbitrary 10-hour threshold
was set for all trawlers and anchovy purse-seiner fleets. Tuna purse-seiners
performed longer trips, so the 10th percentile, i.e. 106 hours, was used as
their threshold. The number of vessels, dyads and the median duration of a dyad
are shown in Table 1.

```{r, echo=FALSE}
library(flextable)
table1 <- read.csv("Table1.csv")
table1$X <- as.character(table1$X)

table_1 <- flextable(
  table1
)
table_1 <- set_header_labels(table_1, X = "",
                             X.1 = "", Cluster.1 = "Cluster 1", 
                             Cluster.2 = "Cluster 2", Cluster.3 = "Cluster 3")
lines_delta <- c(4,9,14,19,24,29)
table_1 <- mk_par(table_1, i = lines_delta, j = 1, value = as_paragraph("(δ = 5km"))
lines_t <- lines_delta + 1
table_1 <- mk_par(table_1, i = lines_t, j = 1, value = as_paragraph("Δt = 1h)"))

library(officer)
std_border = fp_border(color="black")
lines_h <- c(5,10, 15, 20, 25)
table_1 <- hline(table_1, i = lines_h, j = 1:6, border = std_border)
table_1 <- vline(table_1, j = 2, border = std_border, part = "all")
table_1 <- autofit(table_1, add_w = 0, add_h = 0)
table_1 <- set_caption(table_1, caption = "Table 1. Statistics per fleet of number of vessels, number of dyads, their duration (median in hours), the δ threshold for Prox, and the frequency of record transmission. The first three statistics are also displayed for each cluster.")
table_1
```

### Joint movement metrics

The review made by @Joo2018 defined three dimensions of joint movement:
proximity (closeness in space-time), coordination in direction and coordination
in speed. The article evaluated ten metrics used in the literature to assess
joint movement and showed that some metrics were either redundant or inaccurate
for characterising joint movement, some others were better suited to assess
proximity, and others were more sensitive to coordination. Based on that work,
we chose three metrics that were positively evaluated and that -- together --
account for the different aspects of joint movement: 1) the proximity index
(proximity), 2) dynamic interaction in displacement (coordination in speed, and
in displacement when time steps are regularly spaced), and 3) dynamic
interaction in direction (coordination in direction).

The proximity index (Prox) is defined as the proportion of simultaneous fixes
that are spatially close. To define closeness, we needed to fix a distance
threshold $\delta$. For pair trawlers, it is expected that at the very moment of
fishing, vessels working together are separated by less than 1 km from each
other. When they were not fishing, they could still move together but not
necessarily at <1km. Thus, a $5$km threshold was used for this fleet.
We also used a 5km threshold for large bottom otter trawlers to get comparable
results to those of pair trawlers. Anchovy purse-seiners, mid-water, and small
bottom otter trawlers usually perform short and coastal fishing trips, meaning
that vessels would not necessarily move together as a strategy, but could
sometimes coincide in places due to their short coastal movements. For that
reason, we chose a smaller threshold, 3km, for those three fleets. For tuna
purse-seiners, we chose 10km, as it is roughly the limit of visual detection of
neighbouring vessels.

The calculation of the other two metrics did not require an *ad hoc*
parametrization as for Prox. The dynamic interaction in direction ($DI_\theta$)
and in displacement ($DI_{d}$) measured similarity in direction and
speed/displacement, respectively, between simultaneous fixes (i.e. records of
locations) in a dyad. The mathematical definition of each metric is shown in
Table 2.

```{r, echo=FALSE}
table2 <- read.csv("Table2.csv")
# table1$X <- as.character(table1$X)

table_2 <- flextable(
  table2
)
table_2 <- set_header_labels(table_2, Interpretation.for.joint.movement = "Interpretation for joint movement")

font_size <- 10

table_2 <- mk_par(table_2, i = 1, j = 1, value = as_paragraph("Prox = (Σ", 
  as_chunk(x = "t", 
    props = fp_text(vertical.align = "subscript", font.size = font_size)),
  "1{d",
  as_chunk(x = "t", 
    props = fp_text(vertical.align = "subscript", font.size = font_size)),
  "(A,B) < δ})/T"
)
)

table_2 <- mk_par(table_2, i = 3, j = 1, value = as_paragraph("DI", 
  as_chunk(x = "d", 
    props = fp_text(vertical.align = "subscript", font.size = font_size)),
  " = (Σ",
  as_chunk(x = "t", 
    props = fp_text(vertical.align = "subscript", font.size = font_size)),
  "[1 - (|d",
  as_chunk(x = "t,t+1", 
    props = fp_text(vertical.align = "subscript", font.size = font_size)),
  "(A) - d",
  as_chunk(x = "t,t+1", 
    props = fp_text(vertical.align = "subscript", font.size = font_size)),
  "(B)| / (d",
  as_chunk(x = "t,t+1", 
    props = fp_text(vertical.align = "subscript", font.size = font_size)),
  "(A) + d",
  as_chunk(x = "t,t+1", 
    props = fp_text(vertical.align = "subscript", font.size = font_size)),
  "(B)))",
  as_chunk(x = "β", 
    props = fp_text(vertical.align = "superscript", font.size = font_size)),
  "]) / (T-1)"
)
)

table_2 <- mk_par(table_2, i = 5, j = 1, value = as_paragraph("DI", 
  as_chunk(x = "θ",  
    props = fp_text(vertical.align = "subscript", font.size = font_size)),
  " = (Σ",
  as_chunk(x = "t", 
    props = fp_text(vertical.align = "subscript", font.size = font_size)),
  "cos(θ",
  as_chunk(x = "A", 
    props = fp_text(vertical.align = "subscript", font.size = font_size)),
  as_chunk(x = "t", 
    props = fp_text(vertical.align = "subscript", font.size = font_size)),
  "-θ",
  as_chunk(x = "B", 
    props = fp_text(vertical.align = "subscript", font.size = font_size)),
  as_chunk(x = "t", 
    props = fp_text(vertical.align = "subscript", font.size = font_size)),
  ")) / (T-1)"
)
)

table_2 <- mk_par(table_2, i = c(1,3), j = 2, value = as_paragraph("[0,1]")) 

table_2 <- mk_par(table_2, i = 5, j = 2, value = as_paragraph("[-1,1]")) 

table_2 <- merge_at(table_2, i = 3:4, j = 1, part = "body")
                                                              
table_2 <- autofit(table_2, add_w = 0, add_h = 0)
table_2 <- set_caption(table_2, caption = "Table 2. Joint movement metrics")
table_2
```
*Note:*  $A$, $B$: vessels in the dyad; $T$: number of fixes in the dyad; $d_t(A,B)$: distance in km between vessels $A$ and $B$ at $t$-th fixes; $1{ \{\}}$: index function; 
		$\delta$: distance threshold; 
		$d_{t,t+1}(A)$ (resp. $d_{t,t+1}(B)$): displacement of $A$ (resp. $B$) in km between fixes $t$ and $t+1$; $\beta$ is a scaling parameter for which we assume to take the default value of 1 [@Long2013;@Joo2018];
		$\theta_{A_t}$ (resp. $\theta_{B_t}$): heading of vessel $A$ (resp. $B$) at time $t$. 


### Identification of partners at see with Gaussian mixture models

Partner identification was addressed through a probabilistic clustering approach
using GMMs [@biernacki2006model]. In this approach, each dyad $i$ was
characterised by its three dimensional metrics $X_i=(Prox_i, {DI_d}_i, {DI_\theta}_i)$ 
which were assumed to be a realisation of a three-dimensional
normal distribution. The mean vector and the variance matrix of this
distribution depended on the unknown cluster $Z_i$ to which the dyad $i$
belonged. Given a fixed number of clusters $(G)$ and the three metrics, there
were three elements to estimate for each cluster $g$ $(g=1,...,G)$: a
three-dimensional mean $(\mu_g)$, a $3 \times 3$ covariance matrix $(\Sigma_g)$,
and the proportion of the cluster in the observed dyad population $(\pi_g)$.

In this set-up, the probability density function of given metric values $x_i$ of
a dyad $i$ ($\phi(x_i)$) can be expressed as:

$$\phi(x_i)=\sum_{g=1}^{G} \pi_g f_g(x_i,\mu_g,\Sigma_g)$$ where $\pi_g = P(Z_i = g)$ 
and $ f_g(x_i,\mu_g,\Sigma_g)$ is a three-dimensional Gaussian density
function.

The probability of being in cluster $g$ for each dyad $i$ given the observed
metrics, $P(Z_i = g \vert X_i = x_i)$, also called posterior probability, was
obtained as a by-product of the global estimation of the model and is expressed
as follows:

$$P(Z_i=g \vert X_i=x_i) = \frac{ \pi_g
f_g(x_i,\widehat{\mu}_g,\widehat{\Sigma}_g) }{\displaystyle \sum_{k=1}^{G}\pi_k
f_g(x_i,\widehat{\mu}_k,\widehat{\Sigma}_k) }, $$ where $\widehat{\mu}_g$ and
$\widehat{\Sigma}_g$ stand respectively for the estimated mean in cluster $g$
and the corresponding estimated covariance matrix.

In GMMs, the total number of clusters are chosen according to either statistical
selection criteria (mostly likelihood-based) or case-study goals. A
three-component GMM structure, i.e. $G=3$, was chosen in order to obtain higher
discrepancies between two extreme dyadic-behaviour clusters by allowing to have
a cluster in between corresponding to an intermediate behaviour. This pattern
would be consistent with our expectations of joint movement within the pelagic
pair trawler fleet: dyads moving together all along, some others joining each
other at some moments--like fishing operations, and others moving independently
from each other--likely paired with other vessels.

Each covariance matrix $\Sigma_g$ can be expressed as the product of different
components which specify its orientation, shape and volume (see
@biernacki2006model). We chose a general GMM structure of 3 dyadic-behaviour
clusters allowing for the volume, orientation and shape of the clusters to
differ from one another, called Gaussian\_pk\_Lk\_Ck in @biernacki2006model.

The GMMs were fitted to the pelagic pair trawlers dataset, composed of $6457$
dyads. Parameter estimation was achieved via the iterative EM algorithm. Because
EM is known to be sensitive to initial conditions [@dempster1977maximum], we
fitted 30 different GMMs and kept the one that minimised the integrated complete
likelihood criterion, using the Rmixmod package [@Rmixmod] and based on
@biernacki2006model. From the fitted model, henceforth denoted by $GMM_{pair
trawlers}$, we obtained the posterior probability $P(Z_i=g \mid X_i=x_i)$ of
each dyad $i$ to belong to each cluster $g$ given the metric values $x_i$. We
considered that a dyad was classified as part of the cluster $g$ that maximised
the posterior probability $P(Z_i=g \mid X_i=x_i)$. The level of mixture between
pairs of clusters in the final model was quantified as the overlapping volume
between the tri-Gaussian distributions of each cluster. This index ranges
between 0 (no mixture) and 1 full (mixing). High levels of mixture would
indicate that the clusters are difficult to distinguish from each other, making
the classification poorly relevant.

For each cluster, we computed a global average of the Z-scores (i.e. centred and
scaled transformation) of their $(Prox_i, {DI_d}_i, {DI_\theta}_i)$-features,
and ordered them accordingly. Based on the definitions of the metrics
[@Joo2018], the cluster with the highest average was associated to partners at
sea behaviour.

The GMM fitted on pelagic pair trawlers ($GMM_{pair trawlers}$) was then used on
each of the other fleets to classify their dyads, into the three identified
groups. For each dyad $i$ of the other fleets, we computed $P(Z_i=g \mid X_i=x_i)$ 
for $g=\{1,2,3\}$ and assigned the dyad to the most plausible cluster.

Using GMMs provided several advantages compared to other common clustering
algorithms. Since it is a model-based clustering approach, we obtained posterior
probabilities of belonging to each cluster; it is thus a probabilistic
classification instead of a hard classification. The k-means algorithm can
actually be seen as a particular case of a GMM: the former optimizes a loss
function which could be seen as the negative log likelihood of a GMM with
spherical shape and same variance among clusters [@Bishop2006].
The GMM fitted to the pair trawler data allowed for different variances and was
not constrained to spherical structures, thus being more flexible than k-means.
Moreover, the EM algorithm used to estimate the parameters in the GMM runs a
k-mean algorithm to find a suitable starting point.

### Vessel and fleet characterisation

We focused on the first group of each fleet, i.e. partners at sea. Their
relative importance in the fleets were represented by the proportions of vessels
and dyads involved in the cluster. For each fleet, the social relationships
between vessels that engaged at least once in partners at sea behaviour were
visually represented as a social network [@Scott1988; @Jacoby2016]. The elements
of the sociomatrix of the network, i.e. adjacency matrix, represented the number
of partner-at-sea dyads between the vessels ---that had at least one dyad in the
cluster. The Fruchterman and Reingold algorithm was chosen to draw the graph. It
positions the nodes of the graph in the space so that all edges are more or less
equal length and there are as few crossing edges as possible, aiming at an
aesthetic representation [@Fruchterman1991]. The igraph package was used for
this purpose [@igraph2006R].

We identified which and how many vessels were exclusive, i.e. only formed
partners at sea with one vessel throughout the whole period of study. In the
adjacency matrix this corresponded to the rows with 0 everywhere except once. To
assess how exclusive were partnerships at the fleet level, a loyalty index was
defined as the proportion of vessels that showed exclusiveness in partnership.
For this calculation we excluded vessels with only one dyad in the group.

All the R codes for partner-at-sea identification via GMMs and vessel and fleet
characterisation are available at
[https://rociojoo.github.io/partners-at-sea](https://rociojoo.github.io/partners-at-sea)
(doi:10.5281/zenodo.4016377)


# Results

## Pelagic pair trawlers

```{r, echo=FALSE}

table3 <- read.csv("Table3.csv")

table_3 <- flextable(
  table3
)
table_3 <- set_header_labels(table_3, X = "",
                             X.1 = "", Cluster.1 = "Cluster 1", 
                             Cluster.2 = "Cluster 2", Cluster.3 = "Cluster 3")

table_3 <- mk_par(table_3, i = 1, j = 1, value = as_paragraph("π"))
table_3 <- mk_par(table_3, i = 2, j = 1, value = as_paragraph("μ"))
table_3 <- mk_par(table_3, i = 5, j = 1, value = as_paragraph("Σ",
            as_chunk(x = "ii", 
    props = fp_text(vertical.align = "subscript", font.size = font_size))                                                  ))

table_3 <- mk_par(table_3, i = c(3,6), j = 2, value = as_paragraph("DI",
            as_chunk(x = "θ", 
    props = fp_text(vertical.align = "subscript", font.size = font_size))                                                  ))
table_3 <- mk_par(table_3, i = c(4,7), j = 2, value = as_paragraph("DI",
            as_chunk(x = "d", 
    props = fp_text(vertical.align = "subscript", font.size = font_size))                                                  ))


table_3 <- autofit(table_3, add_w = 0, add_h = 0)
table_3 <- set_caption(table_3, caption = "Table 3. Parameter estimates of GMM for pair trawlers")
table_3
```

```{r, echo=FALSE, warning=FALSE}

table4 <- read.csv("Table4.csv")

table_4 <- flextable(
  table4
)
table_4 <- set_header_labels(table_4, X = "",
                             Cluster.1 = "Cluster 1", X.1 = "Cluster 1", X.2 = "Cluster 1",
                             Cluster.2 = "Cluster 2", X.3 = "Cluster 2", X.4 = "Cluster 2",
                             Cluster.3 = "Cluster 3", X.5 = "Cluster 3", X.6 = "Cluster 3")

table_4 <- merge_h(table_4, part = "header")

table_4 <- mk_par(table_4, i = 1, j = c(3,6,9), value = as_paragraph("DI",
            as_chunk(x = "θ", 
    props = fp_text(vertical.align = "subscript", font.size = font_size))                                                  ))
table_4 <- mk_par(table_4, i = 3, j = 1, value = as_paragraph("DI",
            as_chunk(x = "θ", 
    props = fp_text(vertical.align = "subscript", font.size = font_size))                                                  ))

table_4 <- mk_par(table_4, i = 1, j = c(4,7,10), value = as_paragraph("DI",
            as_chunk(x = "d", 
    props = fp_text(vertical.align = "subscript", font.size = font_size))                                                  ))
table_4 <- mk_par(table_4, i = 4, j = 1, value = as_paragraph("DI",
            as_chunk(x = "d", 
    props = fp_text(vertical.align = "subscript", font.size = font_size))                                                  ))


table_4 <- autofit(table_4, add_w = 0, add_h = 0)
table_4 <- set_caption(table_4, caption = "Table 4. Correlations between metrics per cluster obtained from Σ estimates of the GMM for pair trawlers")
table_4
```

After pre-processing, $6457$ dyads were classified with GMMs. The estimated
parameters are shown in Table 3. The correlations between features (Table 4)
were not negligible, which supports the joint use of metrics that evaluate
different aspects of dyadic movement. There was little overlap between cluster 1
and the other two: $1.9 \times 10^{-3}$ and $3.7 \times 10^{-10}$, between
clusters 1 and 2, and 1 and 3, respectively. There was higher overlap ($0.32$)
between clusters 2 and 3. Moreover, most dyads were classified based on high
values of their posteriors ($1.00$, $0.95$, and $0.86$ as median posteriors for
each group, respectively; Fig. 3), and all of them above $0.5$.

The three clusters obtained corresponded to distinct levels of joint movement
(Fig. 1). The first one (purple in Fig. 1) corresponded to high joint movement
in its three dimensions: proximity, coordination in direction and in
speed/displacement. This was the expected pattern for partnership at sea. The
second one (green in Fig. 1) was associated to a lower degree of joint movement
in all dimensions. The third cluster (yellow in Fig. 1) was overall
characterised by low proximity, relatively low coordination in direction, and
low coordination in displacement. In these two metrics, there was a considerable
amount of overlap, with Prox being the metric that made these two groups
distinguishable. The tracks of the most representative dyad of each cluster,
i.e. the one with the largest $P(Z=g \mid X=x)$, are shown in Fig. 2. Animations
of the trajectories and time series related to the three metrics can be found in
<https://rociojoo.github.io/partners-at-sea/>.

In total, $8\%$, $26\%$ and $66\%$ of the examined dyads were classified in the
first, second and third cluster, respectively (Table 1). The examined dyads were
couples of vessel tracks coinciding in a common area at the same time. Not all
pairs of vessels that cross their paths should be necessarily working together.
On the other hand, most of the vessels of the fleet, $56$ ($95\%$), participated
at least once in dyads classified as partners at sea. From them, $46$ had
exclusive partners (Fig. 5), which translated into a $0.82$ loyalty index for
the fleet.

## Dyads from other fleets

In this section, we focused only on the first group, i.e. partners at sea. The
proportion of dyads classified in each cluster is presented in Table 1, and
examples of dyads in each cluster for all fleets can be found in
<https://rociojoo.github.io/partners-at-sea/>, a companion website for the
manuscript.

When using $GMM_{pair trawlers}$ to classify dyads from the other fleets, we
found partners at sea in all of them except for tuna purse seiners. In all the
fleets, the posterior probabilities computed for classification were relatively
high (medians were $>0.65$ and all posteriors were $>0.5$; Fig. 3) showing low
ambiguity for classification in all groups.

For large, small bottom, mid-water otter trawlers and anchovy purse-seiners,
$312$, $93$, $3$ and $568$ dyads were classified as partners at sea,
respectively (Table 1). In all cases, it represented less than $1\%$ of the
examined dyads, showing that vessels in the same area do not always move
together, and when they do, they do not do it in large groups.

We compared the distribution of values of the metrics in the first group between
pelagic pair trawlers and the other fleets (large and small bottom otter
trawlers, and anchovy purse-seiners; Fig. 4). Large bottom otter trawlers showed
the most similar shapes of the distributions to pair trawlers, for all metrics,
though the values of $DI_{d}$ were less skewed to the right than for pair
trawlers. This difference in skewness for $DI_{d}$ was also true for the other
two fleets. Moreover, 'partners at sea' among anchovy purse-seiners took lower
values of all the metrics (more skewed to the left). Since both fleets target
pelagic species, one might have expected to find similar metric values for their
partners at sea. This difference could be related to the different sampling rate
(10 minutes), which allows looking at a finer scale in joint movement, showing
that at this scale it is rather low. It could also be an indicator of a joint
movement that does not occur at a dyadic scale, i.e. a couple of vessels that
decide to move together; if larger groups were moving together, this pattern
would not have necessarily reflected in very high values in the dyadic movement
metrics.

The percentage of vessels engaged in at-sea partnership and their exclusiveness
varied greatly among fleets (Fig. 5). $38$ out of $266$ large bottom otter
trawlers ($14\%$) showed at-sea partnership at least once, and from them, $19$
had exclusive partners (loyalty = $0.54$). A larger percentage of small bottom
otter trawlers engaged in partnership ($26\%$, or $52$ out of $202$). From them,
$38$ had exclusive partners ($35$ with $>1$ dyad; loyalty = $0.74$). Only $4$
out of $70$ mid-water otter trawlers engaged in partnership, which was exclusive
(loyalty = $1$) and only occurred three times. In contrast, $43\%$ of the
anchovy purse-seiners engaged in partnership (or $327$ out of $757$ vessels).
$134$ of these vessels were exclusive ($132$ with $>1$ dyad; loyalty = $0.44$).
Most anchovy purse-seiners showed joint-movement links with large groups of
vessels (Fig. 5d), which would be consistent with the differences in the metrics
distribution (Fig. 4).


# Discussion

In this work, we aimed at identifying partners at sea in different fleets around
the world. We presented a simple heuristic approach to identify them by means of
joint movement metrics [@Joo2018], use of Gaussian mixture modelling, and taking
pelagic pair trawlers as a 'training' dataset.

Partners at sea were identified in all the examined fisheries, except for tuna
purse-seiners. This could be partly explained by the long duration of their
fishing trips and large range of movement. While the trip duration in the other
fleets ranged between less than a day and four days, tuna purse-seiner fishing
trips lasted about 30 or 40 days. Tuna purse-seiners, not bounded to fish
together, showed that there was no strategy involving dyadic joint movement
throughout their whole trips. However, data exploration showed that some vessels
moved together in pairs for parts of their trips (see
<https://rociojoo.github.io/partners-at-sea/> for an example in group 2). The
identification of trip segments associated to joint movement (i.e. redefining a
dyad) was out of the scope of this work, and remains open for future research.

Mid-water and small bottom otter trawlers performed equally in terms of trip
duration and distances covered. However, the mid-water otter trawler dataset
only contained three partners at sea dyads, suggesting that individual
competition could be higher in this fleet, or that working together would bring
them no benefit, which could be due to their smaller fishing zones or the
spatial behaviour of their targeted fish. Compared to mid-water trawlers, a
higher percentage of both small and large bottom otter trawlers participated in
partnerships, showing that this is a strategy used in these fleets, though it
has not been adopted by the majority of the vessels. These three trawler fleets
are composed of vessels that engage in fishing activities (métiers) that target
demersal or benthic species (fish, crustaceans, cephalopods). From empirical
observations, these métiers are likely to require less synchronous collaboration
than pelagic métiers. Instead, the observed partner-at-sea behaviours could have
been shaped by environmental or physical constraints (e.g. currents,
@Gloaguen2016) that the vessels would be facing in the same fishing area at the
same time, rather than a collaborative fishing strategy.

A third of anchovy purse-seiners moved in partnership at least once during the
analysed fishing season. Though the trips had a short duration ($\sim 17$
hours), the sampling rate from these VMS data was very high ($\sim 10$ minutes).
At such resolution, joint movement patterns were identified. In this intensive
and highly dynamic monospecific fishery, these findings are somehow a surprise
that may be worth studying in more detail in the future. The high number of
vessels in this fleet showing joint movement, and the high number of connections
displayed in its social network, makes it appealing to study joint movement in
larger groups for this fleet.

While it was expected to find partnership in pelagic pair trawlers, the degree
of loyalty in this fleet was previously unknown, thus revealing about their
partnership strategies. $82\%$ of the vessels (or fishers) opted for exclusive
partnerships, and the ones who did not, exchanged partners in very reduced
groups. In large and small bottom otter trawlers, the loyalty between vessels
involved in the partner at sea cluster was lower; small bottom otter trawlers
are involved in larger groups (Fig. 5). Non-exclusive partnerships involved even
larger groups in the anchovy purse-seine fleet. These fleets may be revealing
two opposed partnership strategies: exclusiveness, which would involve
commitment or long-term partnership, and opportunism, in which a vessel would
move jointly with another one (or even a group of vessels) without any previous
history or commitment. We did not assess the associations between partnerships
and belonging to a same company, and it could be appealing for future studies to
analyse if this would correspond to a strategy where the ship-owner requires his
fishing masters to work together.

This work represents a first approach into studying joint movement behaviour and
strategies in fisheries. It highlights the fact that not all trajectories can be
considered as independent, an assumption made in most modelling studies (e.g.
using state space models; @Joo2013; @Gloaguen2015a). Furthermore, it could be
appealing to apply this approach to select, from a set of trajectories, those
that do not show any partnership at sea. This could allow computing Catch per
Unit of Effort only drawn from independent fishing operations. It could also be
used to evaluate potential errors in modelling fleet dynamics. For instance, one
could fit state-space models using independent tracks on one hand and using all
the tracks on the other, and compare the goodness of fit of both models --and
simulation results --to evaluate the biases in state estimations linked to the
dependence between vessels.

In this study, we focused on a very specific scale of joint movement, the dyad,
defined as a unit composed of fishing trip segments of two vessels occurring at
the same time and in a common area. Studying the strategies of fleets like the
tuna purse-seiners could benefit from the development of methods to identify
joint movement at smaller scales (e.g. segments of fishing trips). The
computation of Prox for each dyad depended on a fixed distance threshold. Here,
we made an ad hoc choice of the threshold for each fleet. This choice is not
straightforward; more in-depth studies of dyadic movement should focus on
sensitivity analysis and the development of an automatic choice of the
threshold.

We consider this work as a first approach to studying partnership at sea, with
pelagic pair trawlers' joint movement as a starting point. Future studies could
focus on other types of partnership at sea, pairwise or not. In many fisheries,
like the anchovy purse-seine fishery, the characterisation of joint movement in
larger groups could help understanding the scales of collective behaviour in the
fisheries. Besides joint movement, leader/following dynamics would also be worth
exploring (see a brief discussion in @Joo2018). All of these components would
help characterising spatial behaviour patterns, but it would not be enough to
understand the triggers of these behaviours. A next step would be to understand
the associations between joint movement (or following movement) and external
factors such as the spatial aggregation of the targeted species, the direction
of currents, or management and economic policies. Ultimately, understanding and
modelling fisher movement including its collective components will contribute to
better estimations of local exploitation of resources. More realistic movement
models would allow better simulations of fisher spatial behaviour and effort for
different management scenarios, thus improving decisions for management.


# Acknowledgements

The authors would like to thank Youen Vermard and Fabien Forget for useful
feedback on the French fleets operating in the North-East Atlantic Ocean and the
Indian Ocean, respectively. Youen's feedback on data processing was of great
help, as well. We also acknowledge the collaboration of Ob7 – Observatoire des
Ecosystèmes Pélagiques Tropicaux exploités, for the tuna dataset. The tuna data
used in this study were collected through the Data Collection Framework (Reg
2017/1004 and 2016/1251) funded by both IRD and the European Union. We are also
grateful to Emily Walker for codes related to the tuna purse-seine fishery.
Guidance for use of servers and different computers from Olivier Berthele and
Audric Vigier were key in the first stages of this work, when Rocío Joo's work
computer was a mess. Thanks to both of you.

# Authors' contributions

RJ, SM and NB conceived the study. NG gave valuable insights on fishing
behaviour at sea that were key to the study design and interpretation of
results. RJ led the data processing and analysis, with contributions from PM and
JR. MPE suggested and helped implementing the GMM. RJ led the writing of the
manuscript. SM, NB and MPE made major contributions to the manuscript, and NG
and PM made minor contributions to it.

# Data and codes availability statement

The dyads' metrics along with all of the R codes for GMM and computation of the
fleet characteristics are available on Zenodo:
<https://doi.org/10.5281/zenodo.4016377>. The codes can also be viewed from
<https://rociojoo.github.io/partners-at-sea/data-processing-and-analysis.html>.
Due to confidentiality agreements, the raw VMS data cannot be shared.


# Figure captions

Fig. 1. Histograms of the joint movement metrics for the three clusters (in
purple, green and yellow) for pelagic pair trawlers. It should be noted that
only $DI_\theta$ ranges from -1 to 1, while Prox and $DI_d$ take values from 0
to 1.

Fig. 2. The most representative dyadic example of each cluster for the pelagic
pair trawler fleet, with the values of the metrics. The coordinates were
transformed to avoid disclosing information about the vessels, whose identifiers
are not shown either. a: Dyad from cluster 1. $Prox=1$; $DI_\theta = 1$; $DI_d =
0.98$. b: Dyad from cluster 2. $Prox=0.57$; $DI_\theta = 0$; $DI_d = 0.69$. c:
Dyad from cluster 3. $Prox=0.06$; $DI_\theta = -0.07$; $DI_d = 0.24$

Fig. 3. Boxplots of the posterior probabilities $P(Z_i = g| X_i = x_i)$ of each
dyad $i$ classified in each group. a: Pelagic pair trawlers. b: Large bottom
otter trawlers. c: Small bottom otter trawlers. d: Mid-water otter trawlers. e:
Tuna purse-seiners. f: Anchovy purse-seiners.

Fig. 4. Histograms of the joint movement metrics ($Prox$, $DI_{\theta}$, and
$DI_d$, in the left, centre and right columns, respectively) for the first group
or partners at sea, comparing the pelagic pair trawlers (blue) with each of the
other fleets (mustard). The other fleets are, in row order from top to bottom:
large bottom otter trawlers, small bottom otter trawlers and anchovy
purse-seiners. Tuna purse-seiners and mid-water otter trawlers  are not shown as
no dyad and only three dyads, respectively, were associated with partnership. 
<!-- a: -->
<!-- Large bottom otter trawlers - Prox. b: Large bottom otter trawlers - -->
<!-- $DI_{\theta}$. c: Large bottom otter trawlers - $DI_d$. d: Small bottom otter -->
<!-- trawlers - Prox. e: Small bottom otter trawlers - $DI_{\theta}$. f: Small bottom -->
<!-- otter trawlers - $DI_d$. g: Anchovy purse-seiners - Prox. h: Anchovy -->
<!-- purse-seiners $DI_{\theta}$. i: Anchovy purse-seiners - $DI_d$. -->

Fig. 5. Network representation of partnership for the pelagic pair trawlers (a),
small bottom otter trawlers (b), large bottom otter trawlers (c) and anchovy
purse-seiners (d). Tuna purse-seiners and mid-water otter trawlers are not shown
as no dyad and only three dyads, respectively, were associated with partnership.
Within each network, only vessels that engaged in partnership at sea at least
once were represented. The size of the nodes (vessels) are proportional to the
number of times they were involved in partnership. The thickness of the lines
between nodes are proportional to the number of partnerships between both nodes.
